---
title: Full title of the paper (Capitalized)
author:
  - name: Dominik Leutnant
    affil: 1,2,\ddagger,*
    orcid: 0000-0003-3293-2315
  - name: John Doe
    affil: 2, \dagger, \ddagger
affiliation:
  - num: 1
    address: |
      Muenster University of Applied Sciences - 
      Institute for Infrastructure, Water, Resources, Environment
      Correnstr. 25, 48149 Muenster, Germany
    email: leutnant@fh-muenster.de
  - num: 2
    address: |
      Your department
      Street, City, Country
    email: mail@mail.com
# author citation list in chicago format
authorcitation: |
  Leutnant, D.; Doe, J.
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
secondnote: |
  These authors contributed equally to this work.
correspondence: |
  leutnant@fh-muenster.de; Tel.: +XX-000-00-0000.
# document options
journal: notspecified
type: article
status: submit
# front matter
simplesummary: |
  A Simple summary goes here.
abstract: |
  A single paragraph of about 200 words maximum. For research articles, 
  abstracts should give a pertinent overview of the work. We strongly encourage
  authors to use the following style of structured abstracts, but without 
  headings: 1) Background: Place the question addressed in a broad context and
  highlight the purpose of the study; 2) Methods: Describe briefly the main
  methods or treatments applied; 3) Results: Summarize the article's main 
  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
  The abstract should be an objective representation of the article, it must not 
  contain results which are not presented and substantiated in the main text and 
  should not exaggerate the main conclusions.
# back matter
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
funding: |
  Please add: ``This research received no external funding'' or ``This research 
  was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded 
  by XXX''. Check carefully that the details given are accurate and use the 
  standard spelling of funding agency names at 
  \url{https://search.crossref.org/funding}, any errors may affect your future 
  funding.
institutionalreview: |
  In this section, you should add the Institutional Review Board Statement and 
  approval number, if relevant to your study. You might choose to exclude 
  this statement if the study did not require ethical approval. Please note 
  that the Editorial Office might ask you for further information. Please add 
  “The study was conducted in accordance with the Declaration of Helsinki, 
  and approved by the Institutional Review Board (or Ethics Committee) of 
  NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies 
  involving humans. OR “The animal study protocol was approved by the 
  Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE 
  (protocol code XXX and date of approval).” for studies involving animals. 
  OR “Ethical review and approval were waived for this study due to REASON 
  (please provide a detailed justification).” OR “Not applicable” for
   studies not involving humans or animals.
informedconsent: |
  Any research article describing a study involving humans should contain this 
  statement. Please add ``Informed consent was obtained from all subjects 
  involved in the study.'' OR ``Patient consent was waived due to REASON 
  (please provide a detailed justification).'' OR ``Not applicable'' for 
  studies not involving humans. You might also choose to exclude this statement 
  if the study did not involve humans.
  
  Written informed consent for publication must be obtained from participating 
  patients who can be identified (including by the patients themselves). Please 
  state ``Written informed consent has been obtained from the patient(s) to 
  publish this paper'' if applicable.
dataavailability: |
  We encourage all authors of articles published in MDPI journals to share 
  their research data. In this section, please provide details regarding where 
  data supporting reported results can be found, including links to publicly 
  archived datasets analyzed or generated during the study. Where no new data 
  were created, or where data is unavailable due to privacy or ethical 
  re-strictions, a statement is still required. Suggested Data Availability 
  Statements are available in section “MDPI Research Data Policies” at 
  \url{https://www.mdpi.com/ethics}.
conflictsofinterest: |
  Declare conflicts of interest or state 'The authors declare no conflict of 
  interest.' Authors must identify and declare any personal circumstances or
  interest that may be perceived as inappropriately influencing the
  representation or interpretation of reported research results. Any role of the
  funding sponsors in the design of the study; in the collection, analyses or 
  interpretation of data in the writing of the manuscript, or in the decision to 
  publish the results must be declared in this section. If there is no role, 
  please state 'The founding sponsors had no role in the design of the study; 
  in the collection, analyses, or interpretation of data; in the writing of the 
  manuscript, an in the decision to publish the results'.
sampleavailability: |
  Samples of the compounds ...... are available from the authors.
supplementary: |
 The following supporting information can be downloaded at:  
 \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: LD 
    long: linear dichroism
bibliography: mybibfile.bib
appendix: appendix.tex
endnotes: false
output: 
  rticles::mdpi_article:
    extra_dependencies: longtable
---

# Version

This Rmd-skeleton uses the mdpi Latex template published 2023-03-25. 
However, the official template gets more frequently updated than the **rticles**
package. Therefore, please make sure prior to paper submission, that you're 
using the most recent .cls, .tex and .bst files 
(available [here](http://www.mdpi.com/authors/latex)).

# Article Header Information

The YAML header includes information needed mainly for formatting the front and 
back matter of the article. Required elements include:

```yaml
title: Title of the paper
author:
  - name: first and last name
    affil: |
      One or more comma seperated numbers corresponding to affilitation
      and one or more  comma seperated symbols corresponding 
      optional notes.
    orcid: optional orcid number
affiliation:  
  - num: 1,..., n for each affiliation
    address: required
    email: required
authorcitation: |
  Lastname, F.
correspondence: |
  email@email.com; Tel.: +XX-000-00-0000.
journal: notspecified
type: article
status: submit
```

Journal options are in Table \ref{tab:mdpinames}. The `status` variable should 
generally not be changed by authors. The `type` variable describes 
the type of of submission and defaults to `article` but can be replaced with any of the ones in Table \ref{tab:mdpitype}

```{r mdpitype, echo = FALSE}
type <- c("abstract, addendum, article, book, bookreview, briefreport, 
casereport, comment, commentary, communication, conferenceproceedings, 
correction, conferencereport, entry, expressionofconcern, extendedabstract, 
datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, 
obituary, opinion, projectreport, reply, retraction, review, perspective, 
protocol, shortnote, studyprotocol, systematicreview, supfile, 
technicalnote, viewpoint, guidelines, registeredreport, tutorial")
knitr::kable(rticles::string_to_table(type, 3), 
             col.names = NULL, format = "latex", booktab = TRUE, 
             caption = "MDPI article types.")
```

## Journal Specific YAML variables

```yaml
# for journal Diversity,
# add the Life Science Identifier using:
lsid: http://zoobank.org/urn:lsid:zoobank.org:act:nnnn


# for journal Applied Sciences
# add featured application
featuredapplication: |
  Authors are encouraged to provide a concise 
  description of the specific application or 
  a potential application of the work. This 
  section is not mandatory.

# for the journal Data
# add dataset doi and license
dataset: https://doi.org/10.1000/182
datasetlicense: CC-BY-4.0

# for the journal Toxins
# add key contributions
keycontributions: |
  The breakthroughs or highlights of the manuscript. 
  Authors can write one or two sentences to describe 
  the most important part of the paper.

# for the journal Encyclopedia
encyclopediadef: |
  For entry manuscripts only: please provide a brief overview
  of the entry title instead of an abstract.
entrylink: The Link to this entry published on the encyclopedia platform.

# for the journal Advances in Respiratory Medicine
# add highlights
addhighlights: |
  This is an obligatory section in “Advances in Respiratory Medicine”, 
  whose goal is to increase the discoverability and readability of the
  article via search engines and other scholars. Highlights should not 
  be a copy of the abstract, but a simple text allowing the reader to 
  quickly and simplified find out what the article is about and what can 
  be cited from it. Each of these parts should be devoted up to 2~bullet 
  points.

```


```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F}
rm(list=ls())
# CONFIGURACIÓN GENERAL
library(knitr)
options(width = 100)
# Opciones generales chunks
opts_chunk$set(echo=T,message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, tidy = F, cache.path = '.cache/', fig.path = './figure/')

#options(xtable.type = 'html')
knit_hooks$set(inline = function(x) {
  
  if(is.numeric(x)) {
    round(x, getOption('digits'))
  } else {
    paste(as.character(x), collapse = ', ')
  }
})
#knit_hooks$set(plot = knitr:::hook_plot_html)
```


```{r}
library(pacman)
packages = c("MASS","knitr","tidyverse","car",'dplyr','kableExtra',"tidyr","readr")
pacman::p_load(char=packages)
```

# Carga de datos

## Vulnerabilidad
```{r}
vuln <- read_delim("./data/vulnerabilidad.csv", 
    delim = ";", escape_double = FALSE, col_types = cols(`Geo Point` = col_skip(), 
        `Geo Shape` = col_skip()), 
    trim_ws = TRUE)%>%arrange(nombre)

colnames(vuln)[colnames(vuln) == "nombre"] <-"Barrio"

vuln$Barrio<-factor(vuln$Barrio,levels = unique(vuln$Barrio))
```

## Barrios
```{r}
barrios <- read_delim("./data/barris-barrios.csv", 
    delim = ";", escape_double = FALSE, col_types = cols(geo_point_2d = col_skip(), geo_shape = col_skip()), 
    trim_ws = TRUE)%>%arrange(Nombre)

colnames(barrios)[colnames(barrios) == "Nombre"] <-"Barrio"

barrios$Barrio<-factor(barrios$Barrio,levels = unique(barrios$Barrio))
```

## Precio de compra y alquiler
```{r}
p_compra <- read_delim("./data/precio-de-compra-en-idealista.csv", 
    delim = ";", escape_double = FALSE, col_types = cols(`Geo Point` = col_skip(), 
        `Geo Shape` = col_skip(), Fecha_creacion = col_skip()), 
    trim_ws = TRUE)%>%arrange(BARRIO)

p_compra$BARRIO<-factor(p_compra$BARRIO,levels = unique(p_compra$BARRIO))

p_alquiler <- read_delim("./data/precio-alquiler-vivienda.csv", 
    delim = ";", escape_double = FALSE, 
    trim_ws = TRUE)%>%arrange(BARRIO)

p_alquiler$BARRIO<-factor(p_alquiler$BARRIO,levels = unique(p_alquiler$BARRIO))

precios<-full_join(p_compra,p_alquiler,by="BARRIO",suffix = c(" de compra"," de alquiler"))
colnames(precios)[colnames(precios) == "BARRIO"] <-"Barrio"

head(precios)
```

## Recibos IBI
```{r}
ibi <- read_delim("./data/rebuts-ibi-2022.csv", delim = ";", escape_double = FALSE, col_types = cols(ID = col_skip(), geo_shape = col_skip(), geo_point_2d = col_skip(), Año = col_skip()), trim_ws = TRUE)%>%
  arrange(Barrio)%>%
  mutate_at(vars(-all_of(c("Distrito","Barrio"))), as.numeric)

ibi$Barrio<-factor(ibi$Barrio,levels = unique(ibi$Barrio))

# Hacemos la media de las observaciones de cada barrio en los tres año y nos quitamos 2/3 de las observaciones
ibi <- ibi %>% group_by(Barrio) %>%  mutate(across(where(is.numeric), mean, na.rm=TRUE))%>%distinct()
```

## Bancos por barrio
```{r}
bancos <- read_delim("./data/bancs-en-via-publica-bancos-en-via-publica.csv", 
    delim = ";", escape_double = FALSE, col_types = cols(gid = col_skip(), 
        `Num. Policia` = col_skip(), geo_point_2d = col_skip()), 
    trim_ws = TRUE)%>%arrange(Barrio)

bancos$Barrio%<>%gsub("[0-9] - ","",.)
bancos$Barrio%<>%factor(levels = unique(bancos$Barrio))

num_bancos<-bancos%>%group_by(Barrio)%>%summarize(N=n())
head(num_bancos)
```

# Fusionamos todos los dataset

```{r}
df<-barrios%>%full_join(vuln,by="Barrio")%>%full_join(num_bancos,by="Barrio")%>%full_join(precios,by="Barrio")%>%full_join(ibi,by="Barrio")
```

Vemos como hay cuatro barrios más de lo normal, vamos a solucionarlo:

```{r}
tail(df)

# El barrio MONTOLIVET se llama únicamente en el dataset "vuln" y "num_bancos" como MONT-OLIVET
levels(vuln$Barrio)[vuln$Barrio=="MONT-OLIVET"]<-"MONTOLIVET"
levels(num_bancos$Barrio)[num_bancos$Barrio=="MONT-OLIVET"]<-"MONTOLIVET"

# Lo mismo ocurre con la Fonteta de sant lluis y el dataset "ibi
levels(ibi$Barrio)[ibi$Barrio=="FONTETA DE SANT LLUIS"]<-"LA FONTETA S.LLUIS"

# Además, la primera y última observación de barrios no corresponden a ningún barrio, por lo que tenemos que eliminarlas
num_bancos%<>%slice(-c(1,length(num_bancos$N)))
```

Una vez solucionado, volvemos a crear el dataset:

```{r}
df<-barrios%>%full_join(vuln,by="Barrio")%>%full_join(num_bancos,by="Barrio")%>%full_join(precios,by="Barrio")%>%full_join(ibi,by="Barrio")
head(df)
```






